---
title: "Creating Matched Samples for Supermajority Voter Districts"
output:
  html_document:
    theme: readable
    df_print: paged
    highlight: tango
    smart: false
    toc: yes
    toc_float: no
    code_folding: show
---




```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning=F, message=F, fig.width=8)

```






# R Packages


```{r, eval=F}
install.packages( "MatchIt" )
install.packages( "rgenoud" )
install.packages( "ggplot2" )
```


```{r}
library( MatchIt )
library( rgenoud )
library( dplyr )
library( ggplot2 )
library( stargazer )
```



# Load Data

This file was generated from the data steps. 


```{r}
dat <- read.csv( "https://raw.githubusercontent.com/lecy/political-ideology-of-nonprofits/master/DATA/CensusPlusVotingAll.csv", stringsAsFactors=F )
```


```{r, eval=F}
dat <- read.csv( "CensusPlusVotingAll.csv", stringsAsFactors=F )

dat <- 
  dat %>% 
  select( vtd.key1, tract.key, pop10, Shape_area, medianage, 
               income, totalpop, poverty, hispanic, 
               white, black, Pres_D_08, Pres_R_08 ) %>%
  mutate( pop.density = 10000*(pop10 / Shape_area),
          vote.dem = Pres_D_08 / (Pres_D_08+Pres_R_08) )

head( dat )
```





## Restrict Data to Supermajority Voting Districts

Restrict the sample to only districts that voted over 70% in favor of Obama in 2008, or over 70% in favor of McCain.

```{r}
# dat$vote.dem <- dat$Pres_D_08 / (dat$Pres_D_08 + dat$Pres_R_08)

dat$dem.super <- dat$vote.dem >= 0.70 

dat$repub.super <- dat$vote.dem <= 0.30 

match.dat <- dat[ dat$vote.dem <= 0.30 | dat$vote.dem >= 0.70 , ]

match.dat <- na.omit( match.dat )

head( match.dat )
```







# Match Voting Districts

For details on the matching procedures see:

Iacus, S. M., King, G., & Porro, G. (2012). Causal inference without balance checking: Coarsened exact matching. Political analysis, 20(1), 1-24.

Ho, D., Imai, K., King, G., & Stuart, E. A. (2011). MatchIt: Nonparametric Preprocessing for Parametric Causal Inference. Journal of Statistical Software,42(i08).

Sekhon, Jasjeet S. 2011. "Multivariate and Propensity Score Matching Software with Automated Balance Optimization: The Matching package for R." Journal of Statistical Software. 42(7): 1-52.


```{r}
# library( MatchIt )
# library( rgenoud )


my.dat <- match.dat[ c("vtd.key1", "tract.key",
                       "dem.super","pop.density","income",
                       "poverty","black","hispanic","medianage") ]

head( my.dat )

table( my.dat$dem.super ) 

# summary( my.dat )
```


```{r, results="asis"}
match.dat %>%
  select( income, pop.density, medianage, poverty,
          hispanic, white, black ) %>% 
          stargazer( type = "text", digits=0 )
```






## Paper Approach

This approach generates a weighted matched set using a genetic search algorith. Depending upon the underlying data, you can refined the matched sample by selecting pairs of voting districts from each set based upon matching weights. 


```{r, eval=F}

# my.dat$pop.density <- my.dat$pop.density / 1000

m.out <- matchit( dem.super ~ poverty + black + pop.density, data = match.dat, 
                  method = "genetic", discard="both" )

summary( m.out )
summary( match.data(m.out, group = "treat") )
summary( match.data(m.out, group = "control") )



### Save the matched dataset

m.dat <- match.data( m.out )

nrow( m.dat )

length( unique( m.dat$tract.key  ))

table( m.dat$dem.super )



t.test( as.numeric(m.dat$poverty) ~ m.dat$dem.super )
t.test( as.numeric(m.dat$black) ~ m.dat$dem.super )
t.test( m.dat$pop.density ~ m.dat$dem.super )



library( ggplot2 )

ddat <- m.dat

ggplot( ddat, aes( x=poverty, fill=factor(dem.super) )) + 
        geom_density(alpha = 0.5) + # xlim(10,15) +
        xlab( "Poverty Rate" )

ggplot( ddat, aes( x=black, fill=factor(dem.super) )) + 
        geom_density(alpha = 0.5) + # xlim(10,15) +
        xlab( "Percent Black" )

ggplot( ddat, aes( x=pop.density, fill=factor(dem.super) )) + 
        geom_density(alpha = 0.5) + xlim(0,.0025) +
        xlab( "Population Density" )



```




```{r}

match.dat$pop.density <- my.dat$pop.density / 1000

match.dat <- filter( match.dat, pop.density < 30 )



m.out <- matchit( dem.super ~ poverty + black + pop.density, data = match.dat, 
                  method = "genetic", discard="both" )

m.dat <- match.data( m.out )


### CREATE A BALANCED SET WITH ONLY ONE CONTROL COUNTY FOR EACH TREATMENT COUNTY

matches <- m.out$match.matrix

setdiff( row.names( matches),  row.names( match.dat ) )  # all in model dataset

these.matched <- row.names( match.dat ) %in% row.names( matches)



# keep only the closest matches - each treatment group appears once, controls can appear many

dupes.temp <- data.frame( treat=row.names(matches), 
                          control=matches[,1], 
                          distance=m.out$distance[ these.matched ],
                          stringsAsFactors=F )

dupes.temp <- na.omit( dupes.temp[ order(dupes.temp$control) , ] )




table( m.dat$dem.super )

dupes <- NULL

for( i in unique(dupes.temp$control) )
{
    sub.dupes <- dupes.temp[ dupes.temp$control == i , ]
    
    keep <- which( sub.dupes$distance == max(sub.dupes$distance) )
    
    dupes <- rbind( dupes, sub.dupes[ keep, ] )
    
}



these.ones <- c( dupes$treat, dupes$control )

balanced <- match.dat[ row.names(match.dat) %in% these.ones , ]

table( balanced$dem.super )

head( balanced )

t.test( as.numeric(balanced$poverty) ~ balanced$dem.super )
t.test( as.numeric(balanced$black) ~ balanced$dem.super )
t.test( balanced$pop.density ~ balanced$dem.super )


match.dat$pop.density = match.dat$pop.density * 1000
match.dat %>% arrange( - pop.density ) %>% head( 10 )

total.voters <- dat$Pres_D_08 + dat$Pres_R_08
summary( total.voters )



ddat <- match.dat

ggplot( ddat, aes( x=poverty, fill=factor(dem.super) )) + 
        geom_density(alpha = 0.5) + # xlim(10,15) +
        xlab( "Poverty Rate" )

ggplot( ddat, aes( x=black, fill=factor(dem.super) )) + 
        geom_density(alpha = 0.5) + # xlim(10,15) +
        xlab( "Percent Black" )

ggplot( ddat, aes( x=pop.density, fill=factor(dem.super) )) + 
        geom_density(alpha = 0.5) + xlim(0,25) +
        xlab( "Population Density" )



ddat <- balanced

ggplot( ddat, aes( x=poverty, fill=factor(dem.super) )) + 
        geom_density(alpha = 0.5) + # xlim(10,15) +
        xlab( "Poverty Rate" )

ggplot( ddat, aes( x=black, fill=factor(dem.super) )) + 
        geom_density(alpha = 0.5) + # xlim(10,15) +
        xlab( "Percent Black" )

ggplot( ddat, aes( x=pop.density, fill=factor(dem.super) )) + 
        geom_density(alpha = 0.5) + xlim(0,50) +
        xlab( "Population Density" )


```



```{r}
##############


these.treat <- row.names(my.dat) %in% dupes$treat
these.control <- row.names(my.dat) %in% dupes$control

from.to.temp1 <- paste( match.dat[ these.treat , "vtd.key1" ], 
                        "to", match.dat[ dupes$control , "vtd.key1" ], sep="-" )

from.to.temp2 <- paste( match.dat[ these.control , "vtd.key1" ], 
                        "to", match.dat[ these.treat , "vtd.key1" ], sep="-" )

balanced$match.pair <- c( from.to.temp1, from.to.temp2 )




m.out2 <- matchit( dem.super ~ poverty + black + pop.density, data = balanced, 
                  method = "genetic", discard="both", reestimate=TRUE )

summary( m.out2 )
```


## Alternative Approach

```{r}

m.out <- matchit( dem.super ~ poverty + black + pop.density, data = my.dat, 
                  method = "genetic", discard="both", reestimate=TRUE  )

summary( m.out )
summary( match.data(m.out, group = "treat") )
summary( match.data(m.out, group = "control") )

m.dat3 <- match.data( m.out )

summary( m.dat3 )

t.test( poverty ~ dem.super, data=m.dat3 )
t.test( black ~ dem.super, data=m.dat3 )
t.test( hispanic ~ dem.super, data=m.dat3 )
t.test( medianage ~ dem.super, data=m.dat3 )
t.test( pop.density ~ dem.super, data=m.dat3 )



m.out <- matchit( dem.super ~ poverty + hispanic + medianage,
                  data=my.dat, method="exact", discard="both", reestimate=FALSE )

m.out <- matchit( dem.super ~ poverty + hispanic + medianage,
                  data=my.dat, method="exact", discard="both", reestimate=FALSE )

summary( m.out )

m.dat <- match.data( m.out )

summary( match.data(m.out, group = "treat") )
summary( match.data(m.out, group = "control") )


# Refine match by adding population density

m.out.2 <- matchit( dem.super ~ pop.density, data=m.dat, method="genetic", discard="none" )

summary( m.out.2 )

m.dat2 <- match.data( m.out.2, weights="weights2", subclass="subclass2" )

```




## Refine Matched Samples

```{r}

keep.these <- names( table( m.dat$subclass ) ) [table( m.dat$subclass ) == 2 ]
keep.these <- as.numeric( keep.these )

m.dat2 <- m.dat[ m.dat$subclass %in% keep.these , ]

m.dat2 <- m.dat2[ order(m.dat2$subclass,m.dat2$dem.super) , ]

m.dat2 <- m.dat2[ c("geoname","vtd","subclass","dem.super",
                    "income","poverty","black","hispanic","medianage","pop.density")]

m.dat2$pop.density <- round( m.dat2$pop.density, 1)
m.dat2$income <- paste0( "$",format(m.dat2$income,big.mark=","))

names(m.dat2) <- c("census.tract","votind.district","match.group","dem.super",
                   "income","poverty","black","hispanic","medianage","pop.density")

m.dat2

```


## Comparison of Group Means


```{r}

t.test( poverty ~ dem.super, data=m.dat2 )
t.test( black ~ dem.super, data=m.dat2 )
t.test( hispanic ~ dem.super, data=m.dat2 )
t.test( medianage ~ dem.super, data=m.dat2 )
t.test( pop.density ~ dem.super, data=m.dat2 )


```









<br>
<hr>
<br>


# Citation

Lecy, J. D., Ashley, S. R., & Santamarina, F. J. (2019). Do nonprofit missions vary by the political ideology of supporting communities? Some preliminary results. *Public Performance & Management Review*, 42(1), 115-141. [DOWNLOAD](https://github.com/lecy/political-ideology-of-nonprofits/raw/master/assets/Lecy-Ashley-Santamarina-PPMR-2019.pdf)



<br>
<hr>
<br>
<br>


```{css, echo=F, eval=T}
p {
  color: black;
  font-size:1.2em;
  margin: 20px 0 20px 0 !important;
}

p.caption {
  text-align: center;
  font-weight: bold;
}

th { font-weight: bold; }

td {
    padding: 3px 10px 3px 10px !important;
    text-align: center;
}

table
{ 
    margin-left: auto;
    margin-right: auto;
    margin-top:80px;
    margin-bottom:100px;
}

h1, h2, h3{
  margin-top:100px !important;
  margin-bottom:20px !important;
}

h5{
    text-align: center;
    color: gray;
    font-size:0.8em;
}

img {
    max-width: 90%;
    display: block;
    margin-right: auto;
    margin-left: auto;
    margin-top:30px !important;
    margin-bottom:30px !important;
}


.sourceCode {
   margin-top:50px;
}

.pagedtable-wrapper {
   margin-bottom:30px;
}
```

