---
title: "Creating Matched Samples for Supermajority Voter Districts"
output:
  html_document:
    theme: readable
    df_print: paged
    highlight: tango
    smart: false
    toc: yes
    toc_float: no
    code_folding: show
---




```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning=F, message=F, fig.width=8)

```






# R Packages


```{r, eval=F}
install.packages( "MatchIt" )
install.packages( "rgenoud" )
install.packages( "ggplot2" )
```


```{r}
library( MatchIt )    # popular matching package in R
library( rgenoud )    # genetic search used in matching 
library( dplyr )      # data wrangling 
library( ggplot2 )    # nice graphs 
library( stargazer )  # reports descriptive statistics 
library( pander )     # formats tables for printing in RMD files  
```



# Load Data

This file was generated from the data steps. 


```{r, eval=T}
dat <- read.csv( "https://raw.githubusercontent.com/lecy/political-ideology-of-nonprofits/master/DATA/02-processed-data/CensusPlusVotingAll.csv", stringsAsFactors=F )

dat <- 
  dat %>% 
  mutate( pop.density = 10000*(pop10 / Shape_area),
          vote.dem = Pres_D_08 / (Pres_D_08+Pres_R_08) ) %>% 
  select( vote.dem, white, black, poverty, pop.density, 
          medianage, income, totalpop,  hispanic, 
          vtd.key1, tract.key, pop10, Shape_area,
          Pres_D_08, Pres_R_08 ) 

head( dat )
```





# Restrict Data to Supermajority Voting Districts

Restrict the sample to only districts that voted over 70% in favor of Obama in 2008, or over 70% in favor of McCain.

```{r}
dat$dem.super <- dat$vote.dem >= 0.70 

dat$repub.super <- dat$vote.dem <= 0.30 

dat.super <- dat[ dat$vote.dem <= 0.30 | dat$vote.dem >= 0.70 , ]

dat.super <- na.omit( dat.super )

dat.super$party <- ifelse( dat.super$dem.super, "DEM", "REP" )
dat.super$party <- factor( dat.super$party )
```







# Match Voting Districts

```{r, eval=F}
library( MatchIt )
library( rgenoud )
```

For details on the matching procedures see:

Iacus, S. M., King, G., & Porro, G. (2012). Causal inference without balance checking: Coarsened exact matching. Political analysis, 20(1), 1-24.

Ho, D., Imai, K., King, G., & Stuart, E. A. (2011). MatchIt: Nonparametric Preprocessing for Parametric Causal Inference. Journal of Statistical Software,42(i08).

Sekhon, Jasjeet S. 2011. "Multivariate and Propensity Score Matching Software with Automated Balance Optimization: The Matching package for R." Journal of Statistical Software. 42(7): 1-52.




```{r}
match.dat <- dat.super[ c("vtd.key1", "tract.key", "party", "dem.super",
                       "pop.density","income","white",
                       "poverty","black","hispanic","medianage") ]

table( match.dat$party ) 
```

All supermajority districts in the sample: 

```{r, results="asis"}
match.dat %>%
  select( dem.super, poverty, white, black, pop.density, 
          income, medianage, hispanic ) %>% 
          stargazer( type = "html", digits=2 )
```


```{r}
head( match.dat )
```


The supermajority districts have very different demographic profiles. 

```{r}
ddat <- match.dat

ggplot( ddat, aes( x=poverty, fill=party )) + 
        geom_density(alpha = 0.5) + # xlim(10,15) +
        xlab( "Poverty Rate" )

ggplot( ddat, aes( x=black, fill=party )) + 
        geom_density(alpha = 0.5) + # xlim(10,15) +
        xlab( "Percent Black" )

ggplot( ddat, aes( x=pop.density, fill=party )) + 
        geom_density(alpha = 0.5) + xlim(0,25) +
        xlab( "Population Density" )
```





## Matching Algorithm 


We explored several of the matching procedures, but many failed to produce balanced samples.

The approach presented here generates a weighted matched set using a genetic search algorithm available in the **rgenoud** package. 

Depending upon the underlying data, you can refined the matched sample by selecting pairs of voting districts from each set based upon matching weights. 

```{r, results="hide"}
# genetic search is a stochastic algorithm 
# which uses random start points. 
# Use a set seed if you want results to 
# be reproducible. 

set.seed( 1234 )

# reset row names to make subsets easier below

row.names( match.dat ) <- 1:nrow(match.dat) 

m.out <- matchit( dem.super ~ poverty + black + pop.density, 
                  data = match.dat, 
                  method = "genetic", 
                  discard="both", reestimate=TRUE, 
                  replace=FALSE, caliper=.25 )

# discard="both" disregards all cases outside of the common support region 
#   prior to the matching process 
#
# reestimate=TRUE recalculates distance metrics after discarding cases
#   outside of the common support region 
#
# replace=FALSE disallows one "control group" case being matched with multiple 
#   treatment group cases
#
# caliper=0.25 prunes all matches that have a fit worse than a 0.25 sd distance apart

```

If the matching process is successful you will get the following message (suppressed here for brevity because a bunch of output is printed at each iteration):


```
No significant improvement in 4 generations.

Solution Lexical Fitness Value:
9.678960e-03  2.358318e-02  4.812874e-02  6.144835e-02  6.619846e-02  6.355658e-01  9.781473e-01  9.998425e-01  

Parameters at the Solution:

 X[ 1] :	9.991076e+02
 X[ 2] :	4.674203e+02
 X[ 3] :	1.078070e+02
 X[ 4] :	2.048367e+01

Solution Found Generation 9
Number of Generations Run 14

Thu Jul 23 20:45:46 2020
Total run time : 0 hours 0 minutes and 21 seconds
```

The final matched sample is as follows: 

```{r}
m.out$nn %>% pander()
```

The treatment group is democratic supermajority districts, and the control group is republican supermajority districts. 

There were 236 republican districts discarded because they fell outside of the general region of common support, but only 10 democratic districts. This suggests that republican districts tended to be more extreme (very white, more suburban or rural, and much wealthier. 

Of the remaining cases after the discard stage we were able to find "doppenganger" districts for 109 cases or approximately 15% of the remaining candidates:  109 / (109+610) for democrats, and 109 / (109+555) for republicans. 

These are the cases that we would interpret as being demographically "identical" except for political ideology (on a small set of covariates, at least). 


## Examine Matches 

We can check to see whether we have achieved balance:

```{r}
summary( m.out )
```

Graphically: 

```{r, fig.width=10, fig.height=6, results="hide"}
par( mfrow=c(1,2) )
plot( m.out )
plot( m.out, type="jitter" )
```


For some more graphical representations of balance see the [**cobalt package**](https://cran.r-project.org/web/packages/cobalt/vignettes/cobalt_A0_basic_use.html) in R.

We can also print the list of matches. Each number represents the row ID of the observation, and an NA in the control group means no match was found for that particular democratic district. 

```{r}
matches <- data.frame( treatment.group.id = row.names( m.out$match.matrix), 
                       control.group.id = m.out$match.matrix )

head( matches, 25 )
```

When can then examine matched pairs to see how closely they align on the covariates: 

```{r}
# compare quality of matches
# select id (row) of treated and id (row) of comparison,
# compare on model covariates:

# note the nice balance: 

covariates <- c("dem.super","poverty","black","pop.density")

matched.rows <- c(63,889)
match.dat[ matched.rows , covariates ] %>% pander()

matched.rows <- c(64,612)
match.dat[ matched.rows , covariates ] %>% pander()
```


## Save the Balanced Dataset

To extract our new balanced data frame we can use the match.data() function from MatchIt. 

```{r}
balanced.data <- match.data( m.out )
dim( balanced.data )
```

We now have 109 rows for democrats, and 109 rows of their republican dopelganger districts. 

Save the new data frame: 

```{r}
write.csv( balanced.data, "../DATA/02-processed-data/matched-pairs.csv", row.names=F )
```

Although not necessary, we might also save the list of which cases are matched:

```{r}
matches <- na.omit( matches )
write.csv( matches, "../DATA/02-processed-data/matched-pairs-index.csv", row.names=F )
```

It may come in handy if you want to visualize matches, or do some post-hoc sensitivity analysis. 

One other good reason to set the random seed before creating the matched samples. 


## Check Balance

We no longer have any demographic differences between our "treatment" and "control" group. They differ primarily on political ideology. 

```{r}

table( balanced.data$party )


t.test( as.numeric(balanced.data$poverty) ~ balanced.data$party )
t.test( as.numeric(balanced.data$black) ~ balanced.data$party )
t.test( balanced.data$pop.density ~ balanced.data$party )

ddat <- balanced.data

ggplot( ddat, aes( x=poverty, fill=party )) + 
        geom_density(alpha = 0.5) + # xlim(10,15) +
        xlab( "Poverty Rate" )

ggplot( ddat, aes( x=black, fill=party )) + 
        geom_density(alpha = 0.5) + # xlim(10,15) +
        xlab( "Percent Black" )

ggplot( ddat, aes( x=pop.density, fill=party )) + 
        geom_density(alpha = 0.5) + xlim(0,50) +
        xlab( "Population Density" )
```




<br>
<hr>
<br>


# Citation

Lecy, J. D., Ashley, S. R., & Santamarina, F. J. (2019). Do nonprofit missions vary by the political ideology of supporting communities? Some preliminary results. *Public Performance & Management Review*, 42(1), 115-141. [DOWNLOAD](https://github.com/lecy/political-ideology-of-nonprofits/raw/master/assets/Lecy-Ashley-Santamarina-PPMR-2019.pdf)

**Matching Packages:**

Iacus, S. M., King, G., & Porro, G. (2012). Causal inference without balance checking: Coarsened exact matching. Political analysis, 20(1), 1-24.

Ho, D., Imai, K., King, G., & Stuart, E. A. (2011). MatchIt: Nonparametric Preprocessing for Parametric Causal Inference. Journal of Statistical Software,42(i08).

Sekhon, Jasjeet S. 2011. "Multivariate and Propensity Score Matching Software with Automated Balance Optimization: The Matching package for R." Journal of Statistical Software. 42(7): 1-52.


<br>
<hr>
<br>
<br>


```{css, echo=F, eval=T}
p {
  color: black;
  font-size:1.2em;
  margin: 20px 0 20px 0 !important;
}

p.caption {
  text-align: center;
  font-weight: bold;
}

th { font-weight: bold; }

td {
    padding: 3px 10px 3px 10px !important;
    text-align: center;
}

table
{ 
    margin-left: auto;
    margin-right: auto;
    margin-top:80px;
    margin-bottom:100px;
}

h1, h2, h3{
  margin-top:100px !important;
  margin-bottom:20px !important;
}

h5{
    text-align: center;
    color: gray;
    font-size:0.8em;
}

img {
    max-width: 90%;
    display: block;
    margin-right: auto;
    margin-left: auto;
    margin-top:30px !important;
    margin-bottom:30px !important;
}


.sourceCode {
   margin-top:50px;
}

.pagedtable-wrapper {
   margin-bottom:30px;
}
```

